Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-26 concurrency/CONCUR 2023/paper.tex
add a reference
fix breakage

\than{Write more here}

@@

\begin{mathpar}
	\langle \texttt{m}.\ \nodeboxrep\ \texttt{p} \ |\ \treerep\ \texttt{m}\rangle\ \texttt{insert(p, k, v)}\ \langle \nodeboxrep\ \texttt{p}\ |\ \treerep\ (\texttt{m[k} \mapsto \texttt{v])}\rangle

	\langle \texttt{m}.\ \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{m} \rangle\ \texttt{lookup(p, k)}\ \langle \texttt{v}.\ \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{m} \land \texttt{m(k) = v}\rangle

	\langle \texttt{m}.\ \nodeboxrep\ \texttt{p} \ |\ \treerep\ \texttt{m}\rangle\ \texttt{delete(p, k)}\ \langle \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{(m[k} \mapsto \_ \texttt{])}\rangle
\end{mathpar}

@@

\subsection{Templates vs. Internal Reorganization}
\label{internal_reorganize}
Our original goal was to verify an existing concurrent BST implementation with the template approach, refactoring as necessary but otherwise leaving the code unchanged. However, there is a fundamental mismatch between the template approach and the standard approach to deletion in a BST. The operations we have seen thus far can be factored into a data-structure-agnostic concurrent step (\lstinline{traverse}) and a concurrency-unaware data structure step (\lstinline{insert} or \lstinline{lookup}). When removing a node from a BST, on the other hand, we usually restructure the tree with a \lstinline{pushdown} operation, %include figure?
rotating three-node sections of the tree until the node to be deleted is at a leaf. This operation is simultaneously concurrency-aware (we must lock the nodes involved to avoid race conditions) and data-structure-specific (we need to know precisely which nodes to target and how to rearrange them). This contrasts with linked-list deletion, which only involves the target node and its immediate neighbors. In practice, the existing template proofs for tree structures either do not include deletion at all, or use a very lightweight form of deletion in which nodes are marked as ``deleted'' but are never actually removed from the data structure.

>>>>

\than{William, I've already added the reference of a paper we discussed before}

Proving highly-concurrent traversals correct \cite{feldman2020proving}

@@

\begin{mathpar}
	\centering
{\color{blue} \langle \texttt{m}.\ \nodeboxrep\ \texttt{p} \ |\ \treerep\ \texttt{m}\rangle }

\texttt{insert(p, k, v)}

{\color{blue} \langle \nodeboxrep\ \texttt{p}\ |\ \treerep\ (\texttt{m[k} \mapsto \texttt{v])}\rangle}

{\color{blue}	\langle \texttt{m}.\ \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{m} \rangle }

 \texttt{lookup(p, k)}  \

{\color{blue}\langle \texttt{v}.\ \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{m} \land \texttt{m(k) = v}\rangle }

{\color{blue}	\langle \texttt{m}.\ \nodeboxrep\ \texttt{p} \ |\ \treerep\ \texttt{m}\rangle }

\texttt{delete(p, k)} \ \ \ \ \ \ \

{\color{blue} \langle \nodeboxrep\ \texttt{p}\ |\ \treerep\ \texttt{(m[k} \mapsto \_ \texttt{])}\rangle }
\end{mathpar}

@@

\subsection{Templates vs. Internal Reorganization}
\label{internal_reorganize}
Our original goal was to verify an existing concurrent BST implementation with the template approach, refactoring as necessary but otherwise leaving the code unchanged. However, there is a fundamental mismatch between the template approach and the standard approach to deletion in a BST. The operations we have seen thus far can be factored into a data-structure-agnostic concurrent step (\lstinline{traverse}) and a concurrency-unaware data structure step (\lstinline{insert} or \lstinline{lookup}). When removing a node from a BST, on the other hand, we usually restructure the tree with a \lstinline{pushdown_left} operation, %include figure?
rotating three-node sections of the tree until the node to be deleted is at a leaf. This operation is simultaneously concurrency-aware (we must lock the nodes involved to avoid race conditions) and data-structure-specific (we need to know precisely which nodes to target and how to rearrange them). This contrasts with linked-list deletion, which only involves the target node and its immediate neighbors. In practice, the existing template proofs for tree structures either do not include deletion at all, or use a very lightweight form of deletion in which nodes are marked as ``deleted'' but are never actually removed from the data structure.
